---
title: "Tidyverse Practice"
author: "Paulo Magalang"
date: "12/11/2021"
output: html_document
---

Here are the packages you will need to complete these practice examples:
```{r, message = FALSE, warning = FALSE}
library(tidyverse)
```

and the data you will need (NOTE: you need to put the data in a folder called 'data')
```{r, message = FALSE}
cars <- read_csv("data/auto_mpg.csv")
chickenpox <- read_csv("data/hungary_chickenpox.csv")
```

Check to see if the data loaded in properly.
```{r}
head(cars)
head(chickenpox)
```

## Auto-mpg Dataset

The auto-mpg dataset contains information on the mileage of cars as well as a car's features.

Here is some light data preprocessing that I have done for you:
```{r}
colnames(cars)[7] <- "model_year"
colnames(cars)[9] <- "car_name"
```


1. Because we are unfamiliar with these data, we need to do some exploratory analyses first. Generate a bar plot
that describes the number of samples that belong to a car brand. (Hint: you need to generate a summary table using
summarize())

```{r}
sample_counts <- cars %>% group_by(car_name) %>%
  summarize(count = n())

sample_counts %>% ggplot(aes(x = car_name, y = count)) + geom_col() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) + 
  xlab("Car Brand") + ylab("Num. of Samples") + ggtitle("Number of Samples per Car Brand")
```

2. From our bar plot, it seems like there are some car brands that do not have many samples. Filter out
car brands from our dataset that have less than or equal to 3 samples. Save the filtered data as a variable
named 'cars_filtered'

```{r}
# first determine the car brands with low numbers of samples
sample_counts %>% filter(count <= 3)
```

```{r}
# now that we know which brands have low counts, let's filter them out from our entire 'cars' dataset
cars_filtered <- cars %>% filter(car_name != "bmw", car_name != "cadillac", car_name != "capri", car_name != "chevy",
                                 car_name != "hi", car_name != "mercedes", car_name != "nissan", car_name != "triumph")
```


3. With our filtered dataset, let's visualize the mean mpg and mean acceleration as a grouped bar plot.

```{r}
cars_summary <- cars_filtered %>% group_by(car_name) %>%
  summarize(mean_mpg = mean(mpg),
            mean_acc = mean(acceleration)) %>%
  pivot_longer(c(mean_mpg, mean_acc),
               names_to = "car_feature",
               values_to = "value")
```

```{r}
cars_summary %>% ggplot(aes(x = car_feature, y = value, fill = car_name)) + geom_col(position = "dodge") +
  xlab("Car Feature")
```

4. Let's also generate a grouped box plot of acceleration and mpg for our filtered data.

```{r}
cars_filtered %>% select(car_name, mpg, acceleration) %>%
  pivot_longer(c(mpg, acceleration),
               names_to = "car_feature",
               values_to = "value") %>%
  ggplot(aes(x = car_feature, y = value, fill = car_name)) + geom_boxplot()
```


## Hungary chickenpox dataset

This dataset contains information on chickenpox cases from counties in Hungary over 2005 to 2014.


Some data preprocessing that I have done for you:
```{r}
chickenpox <- chickenpox %>% separate(Date, into = c("day", "month", "year"), sep = "/", convert = FALSE)
```


1. Using the data from 2005 only, generate a line plot that shows how cases fluctuate per month.

```{r}
chickenpox_2005 <- chickenpox %>% filter(year == "2005") %>% 
  pivot_longer(-c(day, month, year), names_to = "county", values_to = "cases") %>%
  group_by(month, county) %>% summarize(total_cases = sum(cases))

chickenpox_2005
```

```{r}
chickenpox_2005 %>% ggplot(aes(x = month, y = total_cases, group = county, color = county)) + geom_line()
```


2. Now, make a line plot that shows the number of cases per month averaged over all years (2005 to 2014) for each
county.

```{r}
chickenpox_long <- chickenpox %>%
  pivot_longer(cols = -c(day, month, year), names_to = "county", values_to = "cases")

chickenpox_avg <- chickenpox_long %>% group_by(month, county) %>%
  summarize(avg_cases = mean(cases))

chickenpox_avg %>% ggplot(aes(x = month, y = avg_cases, group = county, color = county)) + geom_line()
```






































